Sketch Recognition using Particle Swarm Optimizations

### 2009-05-31 paper2/Paper2.tex
finish the paper

The second experiment we implemented was to investigate the effect of symbol complexity and its type on the recognition rate. Figure \ref{fig:test2} and \ref{fig:LDtest2} show the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols (such as square, parallegram). The results also indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.  This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} acheives better segmentation. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of curve or lines only the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first into line segments then tries to decide whether each segment can be represented better as a curve. However, in \textsl{AlgS2} the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the symbols. The table shows that errors are only between two sets of symbols: moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This leads us to the next experiment to choose best set of feature to recognize symbols. 

@@

Different feature sets are tested to determine the best features that can be used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4}). Results show that  geometerical features (\textbf{FS1}) and  Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding geometerical structural features \textbf{(FS1)}.

@@

This paper presented a new approach to sketch recognition using PSO. The system uses both speed and curvature information the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results. The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}. The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

>>>>

The second experiment we implemented was to investigate the effect of symbol complexity and its type on the recognition rate. Figure \ref{fig:test2} and \ref{fig:LDtest2} show the achieved accuracy of each symbol by our algorithms compared to (\textsl{Alg3}) \cite{earlyprocess}. It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols (such as square, parallegram). The results also indicate that algorithm \textsl{AlgS1} achieves better performance than algorithm \textsl{AlgS2} in the symbols that consists only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Figure \ref{fig:LDtest2} shows that in logic design (LD-DS) \textsl{AlgS2} achieved better performance than \textsl{Alg3} and  \textsl{AlgS1}.  This proves that for complex shapes that are combination of lines and arcs are \textsl{AlgS2} acheives better segmentation. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of curve or lines only the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first into line segments then tries to decide whether each segment can be represented better as a curve. However, in \textsl{AlgS2} the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix for the (HS-DS) symbols. The table shows that errors occurs mostly between two sets of symbols: moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This leads us to the next experiment to choose the best set of feature to recognize symbols. 

@@

Different feature sets are tested to determine the best features that can be used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of using different feature sets (from the basic four sets \textbf{FS1,FS2,FS3,FS4} in Section \ref{sec:Recognition}). Results show that  geometerical features (\textbf{FS1}) and  Rubine features (\textbf{FS2}) \cite{gestureexample12} gives the low results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results and it is furthur improved by adding geometerical structural features \textbf{(FS1)}.

@@

This paper presented a new approach to sketch recognition using PSO. The system uses both sthe peed and curvature information as input to the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results. The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}. The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.
